noam_data_exploration_v1

### 2024-07-23 noam_data_exploration_v3.Rmd
Added a cover vs cv figure, but I don't like it. Played around with how to calculate portfolio effects some more.

    transect_level_portfolio[i, 2] <- pe_avg_cv(current_matrix) # calculate portfolio effects for current time series
  }

@@

gross_synchrony_transect <- codyn::synchrony(df = transect_level,
                                   time.var = "Year",
                                   species.var = "Taxa",
                                   abundance.var = "Mean_cover",
                                   metric = "Gross",
                                   replicate.var = "Stratum") # lots of warnings in this one - examine further later
colnames(gross_synchrony_transect) = c("Stratum", "Gross_synchrony") # rename cols for merge later

@@

diversity_stability <- merge(diversity_stability, lm_synchrony_transect, by = "Stratum") # add in loreau and de Mazancourt synchrony
diversity_stability <- merge(diversity_stability, gross_synchrony_transect, by = "Stratum")
diversity_stability <- merge(diversity_stability, transect_level_portfolio, by = "Stratum") # add in portfolio effects
pairs(~ Cover_mean + Cover_cv + Cover_stability + 
        Shannon_mean + Shannon_cv + Shannon_stability + 
        Loreau_synchrony + Gross_synchrony + Portfolio_effect, data = diversity_stability)
```

@@

(gross_stability_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +

(portfolio_stability_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Portfolio_effect, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +

@@

(gross_diversity_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Shannon_mean, y = Gross_synchrony, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +

@@

    quadrat_level_portfolio[i, 2] <- pe_avg_cv(current_matrix) # calculate portfolio effects for current time series
  }

@@

gross_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
                                   time.var = "Year",
                                   species.var = "Taxa",
                                   abundance.var = "Percent_cover",
                                   metric = "Gross",
                                   replicate.var = "Location") # lots of warnings in this one - examine further later
colnames(gross_synchrony_transect) = c("Location", "Gross_synchrony") # rename cols for merge later

@@

>>>>

    transect_level_portfolio[i, 2] <- pe_mv(current_matrix, "loess_detrended") # calculate portfolio effects for current time series
  }

@@

# gross_synchrony_transect <- codyn::synchrony(df = transect_level,
#                                    time.var = "Year",
#                                    species.var = "Taxa",
#                                    abundance.var = "Mean_cover",
#                                    metric = "Gross",
#                                    replicate.var = "Stratum") # lots of warnings in this one - examine further later
# colnames(gross_synchrony_transect) = c("Stratum", "Gross_synchrony") # rename cols for merge later

@@

diversity_stability <- merge(diversity_stability, lm_synchrony_transect, by = "Stratum") # add in loreau and de Mazancourt synchrony
# diversity_stability <- merge(diversity_stability, gross_synchrony_transect, by = "Stratum")
diversity_stability <- merge(diversity_stability, transect_level_portfolio, by = "Stratum") # add in portfolio effects
pairs(~ Cover_mean + Cover_cv + Cover_stability + 
        Shannon_mean + Shannon_cv + Shannon_stability + 
        Loreau_synchrony + Portfolio_effect, data = diversity_stability)
```

@@

# (gross_stability_plot = 
#   ggplot(data = diversity_stability) +
#   geom_point(aes(x = Gross_synchrony, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
#    scale_colour_manual(values = site_colours) +
#    scale_shape_manual(values = habitat_shapes) +
#   facet_wrap(~Site) +
#    
#   theme_bw() + 
#   theme(axis.title.x = element_text(size=18), 
#         axis.title.y = element_text(size=18), 
#         panel.background = element_blank(), 
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_blank(),
#         legend.text = element_text(size = rel(1)), # increase legend text
#         legend.title = element_text(size = rel(1)),
#         legend.position = 'bottom', # move legend to bottom
#         legend.box = 'horizontal', # make legend horizontal
#         legend.box.background = element_rect(colour = "black"))  # add box around legend
#          )
# 
(portfolio_stability_plot = 
  ggplot(data = diversity_stability) +
  geom_point(aes(x = Portfolio_effect, y = Cover_stability, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
   scale_colour_manual(values = site_colours) +
   scale_shape_manual(values = habitat_shapes) +
  facet_wrap(~Site) +

@@

# (gross_diversity_plot = 
#   ggplot(data = diversity_stability) +
#   geom_point(aes(x = Shannon_mean, y = Gross_synchrony, colour = Site, shape = Habitat), size = 8, stroke = 2) + 
#    scale_colour_manual(values = site_colours) +
#    scale_shape_manual(values = habitat_shapes) +
#   facet_wrap(~Site) +
#    
#   theme_bw() + 
#   theme(axis.title.x = element_text(size=18), 
#         axis.title.y = element_text(size=18), 
#         panel.background = element_blank(), 
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_blank(),
#         legend.text = element_text(size = rel(1)), # increase legend text
#         legend.title = element_text(size = rel(1)),
#         legend.position = 'bottom', # move legend to bottom
#         legend.box = 'horizontal', # make legend horizontal
#         legend.box.background = element_rect(colour = "black"))  # add box around legend
#          )

@@

    quadrat_level_portfolio[i, 2] <- pe_mv(current_matrix, "loess_detrended") # calculate portfolio effects for current time series
  }

@@

## gross_synchrony_quadrat <- codyn::synchrony(df = long_data_sub,
##                                    time.var = "Year",
##                                    species.var = "Taxa",
##                                    abundance.var = "Percent_cover",
##                                    metric = "Gross",
##                                    replicate.var = "Location") # lots of warnings in this one - examine further later
## colnames(gross_synchrony_transect) = c("Location", "Gross_synchrony") # rename cols for merge later

@@

```{r cover and cv, fig.height = 8, fig.width = 10}
(cover_cv_plot = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Cover_mean, y = Cover_cv, colour = Site, shape = Habitat), size = 4) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
   facet_wrap(~Site) +

(cover_cv_plot = 
  ggplot(data = diversity_stability_quadrat) +
  geom_point(aes(x = Cover_mean, y = Shannon_mean, colour = Site, shape = Habitat), size = 4) + 
   scale_colour_manual(values = site_colours) +
  scale_shape_manual(values = habitat_shapes) +
   facet_wrap(~Site) +
